Thesis Proposal

### 2019-11-19 notes/proposal/proposal_document.tex
modified proposal doc based on Siva suggestions

\begin{abstract}
	In the classical data analysis setting, where we observe i.i.d. point-cloud data sampled from an unknown distribution $P$, there has been recent interest in the design and analysis of procedures which involve the geometry of $P$. Broadly speaking, these procedures either estimate geometric properties such as high-density regions of $P$, or leverage the geometry of $P$ in some downstream statistical learning task. One way to implicitly adapt to the geometry of $P$ is through the use of graph-based spectral methods. The underlying intuition is that when one appropriately constructs a neighborhood graph, the resulting graph Laplacian encodes a great deal of information about the geometry of $P$; for example, one can view its top eigenvectors as (discrete versions of) functions which are smooth with respect to $P$. For this reason, there has been extensive analysis on consistency properties of graph Laplacians, that is, in what ways and with what error they approximate continuum-level operators.

@@

\section{Background.}

\textcolor{red}{TODO:} This is why you are going to read.

\textcolor{red}{TODO:} This is what you are going to read.

@@

Let $X = \{x_1,
\ldots, x_n\}$ be a sample drawn i.i.d.\ from a distribution $\Pbb$ on $\Rd$,
with density~$f$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
\emph{$r$-neighborhood graph} of $X$, an unweighted, undirected graph with
vertices $V=X$, and an edge $(x_i,x_j) \in E$ if and only if $\norm{x_i -
	x_j} \leq r$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $A \in
\Reals^{n \times n}$ the adjacency matrix, with entries $A_{uv} = 1$ if
$(u,v) \in E$ and $0$ otherwise.  We also denote by $D$ the diagonal degree
matrix, with entries $D_{uu} := \sum_{v \in V} A_{uv}$, and by $I$ the $n
\times n$ identity matrix.

Spectral graph theory involves the study of the eigenvalues and eigenvectors of any of following Laplacian matrices:
\begin{equation*}
\textrm{combinatorial:}~ L = D - A,~~ \textrm{random walk:}~ L_{\textrm{rw}} = I - D^{-1}A,~~ \textrm{normalized:}~ L_{\textrm{n}} = D^{1/2}L_{\textrm{rw}}D^{-1/2}.
\end{equation*}

and is broadly applicable to fields as diverse as the theory of randomized algorithms and differential geometry \textcolor{red}{Chung}. When specifically considering Laplacians of neighborhood graphs such as $G_{n,r}$, one well-developed line of work concerns the asymptotic convergence properties of the various graph Laplacians to associated continuum Laplacian operators.

\begin{itemize}
	\item  \textcolor{red}{Define the Laplacian operator}.
	\item Introduce the study of Laplacians on point-cloud data, manifold learning, and give references as to their background. Discuss the study of the convergence of spectra \textcolor{red}{Belkin, Coifman, Garcia-Trillos and Slepcev 1, Garcia-Trillos and Slepcev Spec, von Luxburg, Hein + von Luxburg, Lafon,Koltchinskii and Gine, Gine and Koltchinskii 1}, pointwise convergence of operators \textcolor{red}{Belkin thesis, Lafon thesis, Hein + Audibert, Singer, Belkin + Niyogi08}, and convergence of smoothness functionals \textcolor{red}{Bousquet + Chapelle, Hein07}. 
	\item Discuss the study of other functionals, such as total variation and normalized cut, on graphs \textcolor{red}{(Garcia-Trillos and Slepcev TV, Arias-Castro, Maier + von Luxburg)}
	\item 
\end{itemize}

Why should I care about the asymptotic convergence of graph Laplacians?
\begin{itemize}
	\item The convergence of graph Laplacians, and their spectra, to underlying continuum operators suggests they may be used to perform a variety of desirable tasks in statistical learning. 
	\item Dimension reduction: under the hypothesis that the data lie on, or are concentrated near, a manifold $\mathcal{M} \subset \Reals^d$ of dimension $m < d$, the eigenvectors $v_1,\ldots,v_m$, can be used to embed $x_i \mapsto (v_{i,1},\ldots v_{i,m})$, reducing the dimension from $d$ to $m$ without losing the geometric structure of $\Pbb$. This algorithm is known as Laplacian Eigenmaps (Belkin).
	\item \textcolor{red}{Clustering: under the hypothesis that the distribution $\Pbb$ is supported on disjoint connected components $\mathcal{C}_1,\ldots,\mathcal{C}_k$, the span of the smallest $k$ eigenvectors of the Laplacian operator $L_n$ will approximately equal the span of the functions $\phi_j(x) = \1\set{x \in \mathcal{C}_j}, j = 1,\ldots,k$.} \textcolor{red}{(Belkin + Arias-Castro)}
	\item Semi-supervised and supervised learning: Discuss the study of Laplacians in supervised learning \textcolor{red}{(Bousquet + Chapelle, Lee + Izbicki, GarciaTrillos + Murray, Padilla)} Discuss the study of Laplacians in semi-supervised learning \textcolor{red}{(Belkin + Niyogi, Zhou + Belkin, Slepcev and Thorpe)}. 
\end{itemize}

In this thesis proposal, we pick up a few of these threads, showing how spectral methods can be used to perform the statistical tasks of density clustering and nonparametric hypothesis testing. Before doing so, we provide some background on each of these areas.

@@

Let $X = \set{x_1,\ldots,x_n}$ consist of i.i.d. samples drawn from a distribution $\Pbb$ on $\Reals^d$, with density $f$. The \emph{density clusters} of $f$, defined as the connected components of
the upper level set $\{x \in \Rd : f(x) \geq \lambda\}$ for some $\lambda > 0$
are a central object of interest in the statistical clustering literature, dating
back to the work of \citet{hartigan1981}, who shows that the classical single-linkage hierarchical clustering algorithm is guaranteed to recover density clusters only when $d = 1$. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level-set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010, balakrishnan2013,kpotufe11} consider consistent estimation of the
cluster tree. None of these works, however, study the ability of spectral algorithms to recover density clusters, which will be our focus.

@@

The results in Section~\ref{subsec:neighborhood_graphs_spectra} detailing with spectral convergence of the graph Laplacian operator have straightforward consequences for the consistency of spectral clustering (for instance, see \cite{vonluxburg2008,garciatrillos18}). However, the behavior of the spectra of these continuum operators can in general be hard to grasp. Therefore, further work relating this spectra to the geometry of the distribution $\Pbb$ is of interest. In this spirit, \citet{shi2009,schiebinger2015,garciatrillos19,ariascastro11} examine the ability of spectral algorithms to recover the latent labels in certain geometrically well-conditioned nonparametric mixture models $\Pbb = \sum_{i = 1}^{k} \pi_i \Pbb_i$. These results focus on global rather than local 
methods, and thus impose global rather than local conditions on the nature
of the density. Moreover, they do not in general guarantee recovery of density 
clusters, which is the focus in our work. Perhaps most importantly, these works
rely on general cluster saliency conditions, which implicitly depend on many
distinct geometric aspects of the cluster $\Cset$ under consideration. We will make
this dependence more explicit, and in doing so expose the role each geometric
condition plays in the clustering problem.

@@

In addition to observing i.i.d samples $x_1,\ldots,x_n$ from a distribution $\mathbb{P}$ with support $\mathcal{X} := [0,1]^d \subset \Reals^d$\footnote{For the purposes of this proposal, whenever we discuss hypothesis testing we will assume $\mathcal{X} = [0,1]^d$, i.e. the unit cube with dimension equal to that of the ambient space $\Reals^d$.}, suppose we observe $Y = \set{y_1,\ldots,y_n}$ according to the following regression model:
\begin{equation}
\label{eqn:regression}
y_i = f(x_i) + \varepsilon_i, \quad \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
\end{equation}
We wish to distinguish
\begin{equation*}
\mathbf{H}_0: f = f_0 := 0 \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: f \neq f_0
\end{equation*}
We will evaluate our performance using worst-case risk: for a given function class $\mathcal{H}$ and test function $\phi: \Reals^n \to \set{0,1}$, let
\begin{equation*}
\mathcal{R}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, f \neq f_0} \Ebb_f(1 - \phi).
\end{equation*}
Without further assumptions on $\mathcal{H}$ no test achieves meaningful power uniformly over all $f \in \mathcal{H}$ \citet{janssen2000}; to make the hypothesis testing question well-posed we will therefore restrict the function class $\mathcal{H}$ in two standard ways. One, when considering type II error we will remove a radius-$\epsilon$ ball in the $L_2$ norm around $f_0$ and consider only the remaining functions
\begin{equation*}
\mathcal{H}_{\epsilon} := \set{f \in \mathcal{H}:\norm{f - f_0}_2 > \epsilon}.
\end{equation*}
Otherwise we can always pick a function $f \neq f_0$ in $\mathcal{H}$ which is indistinguishable from $f_0$ given a finite number of samples. Two, we will insist that the functions in $\mathcal{H}$ display some degree of regularity or smoothness; otherwise it may not be possible to consistently estimate the functional $\norm{f - f_0}_2$.

Let $\alpha \in (0,1)$ be a user input specifying a tolerated level of testing error. A typical way to characterize the performance of a test $\phi$ is through the critical radius $\epsilon_n(\phi; \mathcal{H})$, defined as
\begin{equation*}
\epsilon_n(\phi; \mathcal{H}) = \inf\{\epsilon: \mathcal{R}(\phi; \mathcal{H}_{\epsilon}) \leq \alpha\}.
\end{equation*}
We can then measure the difficulty of a given problem through the minimax critical radius $\epsilon_n^{\star}$, formally 
\begin{equation*}
\epsilon_n^{\star}(\mathcal{H}) = \inf \set{\epsilon:\inf_{\phi} \mathcal{R}(\phi; \mathcal{H}_{\epsilon} ) \leq \alpha}.
\end{equation*}
where the infimum is over all Borel measurable functions $\phi: \Reals^n \to \set{0,1}$. We say a test $\phi$ is minimax optimal over a function class $\mathcal{H}$ if $\epsilon_n(\phi; \mathcal{H}) \asymp \epsilon_n^{\star}(\mathcal{H})$, and we call the rate at which $\epsilon_n^{\star}(\mathcal{H}) \to 0$ as a function of $n$ the \emph{minimax testing rate} over $\mathcal{H}$. To concisely state rates, we will use order notation, in which we write $a_n = O(b_n)$ when $a_n \leq c b_n$ for some constant $c$ and all $n \in \mathbb{N}$, $a_n = o(b_n)$ when $a_n/b_n \to 0$ as $n \to \infty$, and $a_n \asymp b_n$ when $c b_n \leq a_n \leq C b_n$ for constants $c$ and $C$ and all $n \in \mathbb{N}$.

\paragraph{Testing over Sobolev classes.}
The first type of function class $\mathcal{H}$ we consider will be balls in Sobolev spaces. We say a Borel measurable function $f: \mathcal{X} \to \Reals$ is in the space $\mathcal{L}^p(\mathcal{X})$ for $1 \leq p < \infty$ if $\int_{\mathcal{X}} \abs{f(x)}^p \,dx < \infty$. Then, the Sobolev space $W_d^{s,2}(\mathcal{X})$ consists of all functions $f \in \mathcal{L}^1(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$(See \citet{evans10} for a formal definition) belongs to $\mathcal{L}^p(\mathcal{X})$. The Sobolev $\{s,2\}$ norm is then 
\begin{equation*}
\norm{f}_{W_d^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
\end{equation*}
and for a given $L > 0$, the corresponding ball is $W_d^{s,2}(\Xset; L) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq L}$. In this work, we will treat $L$ as a fixed, known constant.

Minimax testing rates for the Sobolev regression testing problem in one dimension were first derived by \citet{ingster82} for the idealized \emph{white noise} regression model:
\begin{equation}
\label{eqn:gaussian_white_noise}
\,dY(x) = f(x) \,dx + \frac{\sigma}{\sqrt{n}}\,dW(x)
\end{equation}
where $\,dW(x)$ denotes a Gaussian white noise process on $\mathcal{X} \subset \Reals$ (defined formally in \citet{gine16} Section 1.2) and $\sigma > 0$ is a known constant. In this case, the minimax test rate is $\epsilon_n^{\star}(\mathcal{W}_d^{s,2}(\mathcal{X}, L)) \asymp n^{-2s/(4s + 1)}$ for all values of $s > 0$. \citet{ingster05} extend this to the multivariate setting where $\mathcal{X} \subseteq \Reals^d$, and show that $\epsilon_n^{\star}(\mathcal{W}_d^{s,2}(\mathcal{X}, L)) \asymp n^{-2s/(4s + d)}$, again for all $d \geq 1, s > 0$. The case where we observe data according to the generative model~\eqref{eqn:regression} was treated in \cite{ingster09}, wherein it is shown that the minimax rate remains the same  $\epsilon_n^{\star}(\mathcal{W}_d^{s,2}(\mathcal{X}, L)) \asymp n^{-2s/(4s + d)}$ so long as $4s > d$. When $4s < d$, if we replace the assumption $f \in W_d^{s,2}(\mathcal{X},L)$ by the assumption $\left(\int_{\mathcal{X}} f^4(x) \,dx\right)^{1/4}$, then \citet{guerre02} show that a test $\phi$ based on the statistic $\frac{1}{n}\sum_{i = 1}^{n}y_i^2$ will have testing error on the order of $n^{-1/4})$ . Since $\mathcal{W}^{s,2}$ does not compactly embed into $\mathcal{L}^4$ when $4s < d$ this result does not apply to Sobolev spaces in this regime; indeed it seems likely that uniformly consistent testing over these spaces is not possible, as the function $f^2$ may no longer have finite variance.

@@

\section{Density clustering with PPR.}

@@

We define test statistics for both the regression and density testing setups. In the former, we show in some preliminary work that our proposed test statistic is minimax optimal over the Sobolev ball $W^{1,2}(\mathcal{X};1)$. In the latter, we empirically demonstrate that our proposed approach has merit. We conclude this section by discussing some practical and computational considerations in this problem. Throughout, we detail the work we intend to complete for a final thesis.

>>>>

\begin{abstract}
	In the classical data analysis setting, where we observe i.i.d. point-cloud data sampled from an unknown distribution $\Pbb$, there has been recent interest in the design and analysis of procedures which involve the geometry of $\Pbb$. Broadly speaking, these procedures either estimate geometric properties such as high-density regions of $\Pbb$, or leverage the geometry of $\Pbb$ in some downstream statistical learning task. One way to implicitly adapt to the geometry of $\Pbb$ is through the use of graph-based spectral methods. The underlying intuition is that when one appropriately constructs a neighborhood graph, the resulting graph Laplacian encodes a great deal of information about the geometry of $\Pbb$; for example, one can view its top eigenvectors as (discrete versions of) functions which are smooth with respect to $\Pbb$. For this reason, there has been extensive analysis on consistency properties of graph Laplacians, that is, in what ways and with what error they approximate continuum-level operators.

@@

\section{Introduction and background.}

This document details the work I intend to complete for my thesis, and is separated into three sections. In the first section, I give background on the convergence properties of neighborhood graphs, as well as the statistical learning problems of clustering and hypothesis testing. In the second section, I detail work I have already completed on clustering using the local spectral algorithm PPR over neighborhood graphs. In the third section, I propose several statistics used to test hypotheses using the spectra of neighborhood graphs, and make explicit the ways in which I will investigate and characterize their behavior. Finally, I give an estimated timeline for completing my proposed work.

@@

Let $X = \{x_1,
\ldots, x_n\}$ be a sample drawn i.i.d. from a distribution $\Pbb$ on $\Rd$,
with density~$f$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
\emph{$r$-neighborhood graph} of $X$, an unweighted, undirected graph with
vertices $V=X$, and an edge $(x_i,x_j) \in E$ if and only if $\norm{x_i -
	x_j} \leq r$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $A \in
\Reals^{n \times n}$ the adjacency matrix, with entries $A_{uv} = 1$ if
$(u,v) \in E$ and $0$ otherwise.  We also denote by $D$ the diagonal degree
matrix, with entries $D_{uu} := \sum_{v \in V} A_{uv}$, and by $I$ the $n
\times n$ identity matrix.

Spectral graph theory involves the study of the eigenvalues and eigenvectors of the graph Laplacian $L := D - A$\footnote{Also the normalized Laplacian matrices $L_{\textrm{rw}} = I - D^{-1}A$ and $L_{\textrm{n}} = D^{1/2}L_{\textrm{rw}}D^{-1/2}$.} and is broadly applicable to fields as diverse as the theory of randomized algorithms and differential geometry \citet{chung97}.

When specifically considering Laplacians of neighborhood graphs such as $G_{n,r}$, one well-developed line of work concerns the asymptotic convergence properties of the various graph Laplacians to associated continuum Laplace operators. Let $\mathcal{X} \subseteq \Reals^d$ be an open, bounded, and connected set with Lipschitz boundary. The Laplace operator $\Delta: C^2(\mathcal{X}) \to L^2(\mathcal{X})$ is defined as
\begin{equation*}
\Delta g = - \dive(\nabla g) =: \sum_{i = 1}^{d} \frac{\partial^2 g}{\partial x_i^2}.
\end{equation*}
There are several ways in which the matrix $L$ can be said to be converging to the operator $\Delta$. The first of these is \emph{functional} convergence. Here, we think of $L$ as acting on functions in $\Reals^d$,
\begin{equation*}
Lg(x) := \frac{1}{n^2r_n^{d+2}}\sum_{i = 1}^{n}(g(x) - g(x_i))A_{ij}
\end{equation*}
and ask whether $Lg(x) \to \Delta g(x)$ in probability as $n \to \infty$ and $r_n \to 0$.  \citet{belkin03,belkin05} consider a slightly different neighborhood graph than $G_{n,r}$ (constructed with a Gaussian rather than uniform kernel) and show that when $\Pbb$ is the uniform distribution over $\mathcal{X}$, such a functional convergence occurs. \footnote{For this and many of the results summarized in this section, $\mathcal{X}$ is permitted to be a low-dimensional manifold, and $\Delta$ is then taken to be the Laplace-Beltrami operator.} \citet{singer06,gine06} extend these results by proving faster convergence rates and a uniform CLT, respectively. When $\Pbb$ is non-uniform the story is slightly different. Now the relevant continuum operator is a weighted analogue of the Laplace operator,
\begin{equation*}
\Delta_{\Pbb} g := - \frac{1}{f} \dive(f^2 \nabla g);
\end{equation*}
\citet{lafon04,hein05} prove the functional convergence $Lg(x) \to \Delta_{\Pbb} g(x)$ in probability as $n \to \infty$ as $n \to \infty$ and $r_n \to 0$.

Another mode of convergence is \emph{smoothness functional} convergence. In this case, we denote $\theta = (g(x_1),\ldots,g(x_n)) \in \Reals^n$, consider the smoothness functional
\begin{equation*}
S(g;\Pbb) := \int_{\mathcal{X}} \norm{\nabla g}_2^2~ f^2(x) \,dx
\end{equation*}
and ask whether $\frac{1}{n^2r_n^{d+2}}\theta^T L \theta \to S(g;\Pbb)$ in probability as $n \to \infty$ and $r_n \to 0$. \citet{bousquet03} proves that the smoothness functional converges pointwise--meaning for a fixed $g \in C^2(\mathcal{X})$--and \citet{hein06} extends this to show uniform convergence over a suitably chosen Holder ball.

The final mode of convergence we mention is \emph{spectral} convergence.  $\Delta_{\Pbb}$ is a positive semidefinite self-adjoint operator, and therefore has a discrete spectrum. Label the $i$th smallest eigenvalue of $\Delta_{\Pbb}$ as $\lambda_i$, with associated eigenfunction $u_i$; similarly, label the $i$th smallest eigenvalue of $L$ as $s_i$ with associated eigenvector $v_i$. We say $L$ converges spectrally to $\Delta_{\Pbb}$ if for any fixed $k \in \mathbb{N}$,
\begin{equation*}
\lim_{n \to \infty} s_{k} = \lambda_k, \quad \textrm{and} \quad \lim_{n \to \infty} \frac{1}{n}\sum_{i = 1}^{n} (u_k(x_i) - v_{k,i})^2 = 0
\end{equation*}
where the limits are taken in probability. \citet{vonluxburg2008} consider the case of a fixed $r_n = r$, and prove spectral convergence of $L$ to an appropriately defined integral operator, \citet{belkin07} prove that when $\Pbb$ is the uniform distribution over $\mathcal{X}$, $L$ converges spectrally to $\Delta_{\Pbb} = \Delta$ for some $r_n \to 0$ as $n \to \infty$, and \citet{garciatrillos18} extend this result to hold for general $\Pbb$ and obtain rates at which $r_n$ may converge to $0$.

The convergence of graph Laplacians to underlying continuum differential operators suggests they may be used to perform a variety of desirable tasks in statistical learning. In this thesis proposal, we study a few of these tasks in detail, and show how spectral methods can be used to perform the statistical tasks of density clustering and nonparametric hypothesis testing. Before doing so, we provide some background on each of these areas.

%\begin{itemize}
	%\item Dimension reduction: under the hypothesis that the data lie on, or are concentrated near, a manifold $\mathcal{M} \subset \Reals^d$ of dimension $m < d$, the eigenvectors $v_1,\ldots,v_m$, can be used to embed $x_i \mapsto (v_{i,1},\ldots v_{i,m})$, reducing the dimension from $d$ to $m$ without losing the geometric structure of $\Pbb$. This algorithm is known as Laplacian Eigenmaps \citet{belkin03a}. 
	%\item Clustering: under the hypothesis that the distribution $\Pbb$ is approximately supported on disjoint connected components $\mathcal{C}_1,\ldots,\mathcal{C}_k$, the embedding $x_i \mapsto (v_{i,1},\ldots,v_{i,k})$ will reveal the cluster structure of $\Pbb$. Subsequent application of a simple clustering method such as $k$-means clustering will then correctly partition the data $X$ into clusters \cite{ng01,ariascastro11,shi2009}.
	%\item Semi-supervised and supervised learning: Discuss the study of Laplacians in supervised learning \textcolor{red}{(Bousquet + Chapelle, Lee + Izbicki, GarciaTrillos + Murray, Padilla)} Discuss the study of Laplacians in semi-supervised learning \textcolor{red}{(Belkin + Niyogi, Zhou + Belkin, Slepcev and Thorpe)}. 
%\end{itemize}

@@

Let $X = \set{x_1,\ldots,x_n}$ consist of i.i.d. samples drawn from a distribution $\Pbb$ on $\Reals^d$, with density $f$. The \emph{density clusters} of $f$, defined as the connected components of
the upper level set $\{x \in \Rd : f(x) \geq \lambda\}$ for some $\lambda > 0$
are a central object of interest in the statistical clustering literature, dating
back to the work of \citet{hartigan1981}, who shows that the classical single-linkage hierarchical clustering algorithm is guaranteed to recover density clusters only when $d = 1$. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level-set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010, balakrishnan2013,kpotufe11} consider consistent estimation of the
cluster tree. None of these works, however, study the ability of spectral algorithms to recover density clusters, which will be our focus in Section~\ref{sec:density_clustering_PPR}.

@@

The results in Section~\ref{subsec:neighborhood_graphs_spectra} have straightforward consequences for the consistency of spectral clustering (for instance, see \cite{vonluxburg2008,garciatrillos18}). However, the behavior of the spectra of these continuum operators can in general be hard to grasp. Therefore, further work relating this spectra to the geometry of the distribution $\Pbb$ is of interest. In this spirit, \citet{shi2009,schiebinger2015,garciatrillos19,ariascastro11} examine the ability of spectral algorithms to recover the latent labels in certain geometrically well-conditioned nonparametric mixture models $\Pbb = \sum_{i = 1}^{k} \pi_i \Pbb_i$. These results focus on global rather than local 
methods, and thus impose global rather than local conditions on the nature
of the density. Moreover, they do not in general guarantee recovery of density 
clusters, which is the focus in our work. Perhaps most importantly, these works
rely on general cluster saliency conditions, which implicitly depend on many
distinct geometric aspects of the cluster $\Cset$ under consideration. We will make
this dependence more explicit, and in doing so expose the role each geometric
condition plays in the clustering problem.

@@

In addition to observing i.i.d samples $x_1,\ldots,x_n$ from a distribution $\mathbb{P}$ with density $p$ supported on $\mathcal{X} := [0,1]^d \subset \Reals^d$\footnote{For the purposes of this proposal, whenever we discuss hypothesis testing we will assume $\mathcal{X} = [0,1]^d$, i.e. the unit cube with dimension equal to that of the ambient space $\Reals^d$.}, suppose we observe $Y = \set{y_1,\ldots,y_n}$ according to the following regression model:
\begin{equation}
\label{eqn:regression}
y_i = f(x_i) + \varepsilon_i, \quad \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1)
\end{equation}
We wish to distinguish
\begin{equation*}
\mathbf{H}_0: f = f_0 := 0 \quad \textrm{vs} \quad \mathbf{H}_{\textrm{a}}: f \neq f_0
\end{equation*}
We will evaluate our performance using worst-case risk: for a given function class $\mathcal{H}$ and test function $\phi: \Reals^n \to \set{0,1}$, let
\begin{equation*}
\mathcal{R}(\phi; \mathcal{H}) = \Ebb_{f = f_0}(\phi) + \sup_{f \in \mathcal{H}, f \neq f_0} \Ebb_f(1 - \phi).
\end{equation*}
Without further assumptions on $\mathcal{H}$ no test achieves meaningful power uniformly over all $f \in \mathcal{H}$ \citet{janssen2000}; to make the hypothesis testing question well-posed we will therefore restrict the function class $\mathcal{H}$ in two standard ways. One, when considering type II error we will remove a radius-$\epsilon$ ball in the $L_2$ norm around $f_0$ and consider only the remaining functions
\begin{equation*}
\mathcal{H}_{\epsilon} := \set{f \in \mathcal{H}:\norm{f - f_0}_2 > \epsilon}.
\end{equation*}
Otherwise we can always pick a function $f \neq f_0$ in $\mathcal{H}$ which is indistinguishable from $f_0$ given a finite number of samples. Two, we will insist that the functions in $\mathcal{H}$ display some degree of regularity or smoothness; otherwise it may not be possible to consistently estimate the functional $\norm{f - f_0}_2$.

Let $\alpha \in (0,1)$ be a user input specifying a tolerated level of testing error. A typical way to characterize the performance of a test $\phi$ is through the critical radius $\epsilon_n(\phi; \mathcal{H})$, defined as
\begin{equation*}
\epsilon_n(\phi; \mathcal{H}) = \inf\{\epsilon: \mathcal{R}(\phi; \mathcal{H}_{\epsilon}) \leq \alpha\}.
\end{equation*}
We can then measure the difficulty of a given problem through the minimax critical radius $\epsilon_n^{\star}$, formally 
\begin{equation*}
\epsilon_n^{\star}(\mathcal{H}) = \inf \set{\epsilon:\inf_{\phi} \mathcal{R}(\phi; \mathcal{H}_{\epsilon} ) \leq \alpha},
\end{equation*}
where the infimum is over all Borel measurable functions $\phi: \Reals^n \to \set{0,1}$. We say a test $\phi$ is minimax optimal over a function class $\mathcal{H}$ if $\epsilon_n(\phi; \mathcal{H}) \asymp \epsilon_n^{\star}(\mathcal{H})$, and we call the rate at which $\epsilon_n^{\star}(\mathcal{H}) \to 0$ as a function of $n$ the \emph{minimax testing rate} over $\mathcal{H}$. To concisely state rates, we will use order notation, in which we write $a_n = O(b_n)$ when $a_n \leq c b_n$ for some constant $c$ and all $n \in \mathbb{N}$, $a_n = o(b_n)$ when $a_n/b_n \to 0$ as $n \to \infty$, and $a_n \asymp b_n$ when $c b_n \leq a_n \leq C b_n$ for constants $c$ and $C$ and all $n \in \mathbb{N}$.

\paragraph{Testing over Sobolev spaces.}
The first type of function class $\mathcal{H}$ we consider will be balls in Sobolev spaces. We say a Borel measurable function $f: \mathcal{X} \to \Reals$ is in the space $\mathcal{L}^p(\mathcal{X})$ for $1 \leq p < \infty$ if 
$\norm{f}_{\mathcal{L}^p(\mathcal{X})} := \int_{\mathcal{X}} \abs{f(x)}^p \,dx < \infty$. Then, the Sobolev space $W_d^{s,2}(\mathcal{X})$ consists of all functions $f \in \mathcal{L}^2(\mathcal{X})$ such that for each $\alpha = (\alpha_1,\ldots,\alpha_d)$ with $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ (See \citet{evans10} for a formal definition) belongs to $\mathcal{L}^2(\mathcal{X})$. The Sobolev $\{s,2\}$ norm is then 
\begin{equation*}
\norm{f}_{W_d^{s,2}(\mathcal{X})}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^2 \,dx
\end{equation*}
and for a given $L > 0$, the corresponding ball is $W_d^{s,2}(\Xset; L) = \set{f: \norm{f}_{W^{s,2}(\Xset)} \leq L}$. (In this work, we will treat $L$ as a fixed, known constant.)

Minimax testing rates for the Sobolev regression testing problem in one dimension were first derived by \citet{ingster82} for the idealized \emph{white noise} regression model:
\begin{equation}
\label{eqn:gaussian_white_noise}
\,dY(x) = f(x) \,dx + \frac{\sigma}{\sqrt{n}}\,dW(x)
\end{equation}
where $\,dW(x)$ denotes a Gaussian white noise process on $\mathcal{X} \subset \Reals$ (defined formally in \citet{gine16} Section 1.2) and $\sigma > 0$ is a known constant. In this case, the minimax test rate is $\epsilon_n^{\star}(\mathcal{W}_d^{s,2}(\mathcal{X}, L)) \asymp n^{-2s/(4s + 1)}$ for all values of $s > 0$. \citet{ingster05} extend this to the multivariate setting where $\mathcal{X} \subseteq \Reals^d$, and show that $\epsilon_n^{\star}(\mathcal{W}_d^{s,2}(\mathcal{X}, L)) \asymp n^{-2s/(4s + d)}$, again for all $d \geq 1, s > 0$. The case where we observe data according to the generative model~\eqref{eqn:regression} was treated in \cite{ingster09}, wherein it is shown that the minimax rate remains the same  $\epsilon_n^{\star}(\mathcal{W}_d^{s,2}(\mathcal{X}, L)) \asymp n^{-2s/(4s + d)}$ so long as $4s > d$. When $4s < d$, if we replace the assumption $f \in W_d^{s,2}(\mathcal{X},L)$ by the assumption $\left(\int_{\mathcal{X}} f^4(x) \,dx\right)^{1/4}$, then \citet{guerre02} show that a test $\phi$ based on the statistic $\frac{1}{n}\sum_{i = 1}^{n}y_i^2$ will have testing error on the order of $n^{-1/4}$ . Since $\mathcal{W}^{s,2}$ does not compactly embed into $\mathcal{L}^4$ when $4s < d$ this result does not apply to Sobolev spaces in this regime; indeed it seems likely that no uniformly consistent test exists over these spaces, as the function $f^2$ may no longer have finite variance.

\paragraph{Testing over Total Variation (TV) spaces.}

To one extent or another, functions in the Sobolev balls exhibit homogeneous smoothness. In reality we may not wish to restrict our attention to such functions. We therefore consider functions of bounded total variation. Formally, for a function $f \in L^1(\mathcal{X})$ the \emph{total variation} semi-norm of $f$ is \citet{evans15}
\begin{equation*}
TV(f;\mathcal{X}) := \sup \left\{ \int_{\mathcal{X}} f \, \dive \, \psi \,dx : \psi \in C_c^1(\mathcal{X}; \Reals^d), \abs{\psi} \leq 1 \right\};
\end{equation*}
and we write $BV_d(\mathcal{X})$ for the subset of functions $f \in L^1(\mathcal{X})$ which have bounded norm
\begin{equation*}
\norm{f}_{BV_d(\mathcal{X})} := \norm{f}_{\infty} + TV(f;\mathcal{X}).
\end{equation*}

In the one-dimensional case, some work has been done on nonparametric testing over the Besov spaces $B_{p,q}^s(\mathcal{X})$ (for a formal definition see Section 4.3 of \cite{gine16}), which we can relate to the bounded variation space $BV_1(\mathcal{X})$ via the embeddings
\begin{equation}
\label{eqn:besov_embeddings}
B_{11}^1(\mathcal{X}) \subseteq BV_1(\mathcal{X}) \subseteq B_{1\infty}^1(\mathcal{X}).
\end{equation}

When $d \geq 2$ minimax testing rates over the Besov and bounded variation spaces are still unknown. However, study of the equivalent estimation problem suggests the answers may be substantially different from the 1d problem. \citet{sadhanala16} consider a discretized version of this problem -- where observations are made on a discrete grid and the regression function is assumed to lie within a discrete bounded variation class -- and show that the minimax estimation rate in empirical norm is on the order of $n^{-1/d}$ for $d \geq 2$. This is slower than minimax estimation rate over the Sobolev space $\mathcal{W}_d^{1,2}$, the largest Sobolev space inscribed within $BV_d$. \citet{delyon96,kerkyacharian08,lepski2015,ruiz18} all consider the regression white noise model \eqref{eqn:gaussian_white_noise} and show, in progressive degrees of generality, that the minimax rates over the Besov spaces $B_{p,q}^s$ exhibit a phase transition at the boundary $2s = d$. When $2s > d$, the minimax estimation rate (with loss measured in $\mathcal{L}^2$ norm) is the standard $n^{-s/(2s + d)}$; on the other hand when $2s < d$, the minimax rate is $n^{-s/2d}$ which is much slower. When $s = 1$, this phase transition occurs at $d = 2$, and so in a minimax sense the univariate and multivariate estimation problems over $BV_d$ are revealed to be quite different. \citet{sadhanala17} show that this same phenomenon carries over to the grid, and the discrete TV classes defined upon it.

@@

\section{Density clustering with PPR.}
\label{sec:density_clustering_PPR}

@@

We define test statistics for both the regression and density testing setups. In the former, we show in some preliminary work that our proposed test statistic is minimax optimal over the Sobolev ball $W^{1,2}(\mathcal{X};1)$. In the latter, we empirically demonstrate that our proposed approach has merit. We conclude this section by discussing some practical and computational considerations for this problem. Throughout, in subsections entitled ``Proposed Work'' we detail the work we intend to complete for a final thesis.
